---
title: "Longitudinal models"
format: html
---

Questions: 
(from Nilam)

1. Is the mean of the latent “language ability” factor scores different across 1st assessment, 2nd assessment, 3rd assessment, etc? This is basically an ANOVA on the latent factor scores – using “assessment number” as the categorical predictor. – an examination of “practice effects”. 

Is the mean of the latent “language ability” factor scores increasing with age?
- This is basically a regression on the latent factor scores – and a growth model.
- In which case we also implicitly test research questions about the variances
- Is the (between-person) variance of the latent “language ability” factor scores increasing with age?

- A more nuanced version of the “practice effects" examination might use delta_t. 
- Is the mean of the latent “language ability” factor different across retest interval?
- This is basically a regression on the latent factor scores.

- In which case we also implicitly test research questions about the variances
- Is the (between-person) variance of the latent “language ability” factor scores increasing with age?

 
3. If we are interested in the covariances –
- Is the structure of the latent “language ability” factor changing with age?
- This again gets us into a multi-group SEM –
- Testing whether the factor loading structure differs across groups.
- This is most easily invoked with using categorical “time variable” – e.g, assessment number, age-group.
- However, it is also possible to do using a continuous time variable (e.g., age). In this case we basically use a moving window to obtain the covariances and then basically use age as a moderator of the factor loadings to examine the age x factor loading interaction.



# Descriptives

## Longitudinal data

```{r}
longitudinal <- d_sub |>
  group_by(dataset_name, subject_id) |>
  count() |>
  filter(n > 1)

d_sub_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(administration_id) |>
  mutate(admin_num = 1:n(), 
         delta_t = c(0,diff(age)))

ggplot(d_sub_long, 
       aes(x = delta_t, fill = as.factor(admin_num))) +
  geom_histogram() + 
  facet_wrap(~dataset_name)
```
## Datasets with CDIs

Longitudinal data with CDI is just FMW2013 right now. 

```{r}
fmw <- filter(d_sub_long, dataset_name == "fmw_2013") |>
  mutate(t = ifelse(age < 23, "t1","t2")) |>
  group_by(subject_id, t) |>
  summarise(age = mean(age), 
            log_rt = mean(log_rt), 
            prod = mean(prod), 
            long_window_accuracy = mean(long_window_accuracy),
            delta_t = mean(delta_t)) |>
  pivot_wider(names_from = "t", 
              values_from = c("age","log_rt","prod","long_window_accuracy", "delta_t"))

summary(lm(prod_t2 ~ long_window_accuracy_t1 * log_rt_t1 + age_t2, 
           data = fmw))
```

## Without CDIs

```{r}
ggplot(d_sub_long, 
       aes(x = age, y = long_window_accuracy, group = subject_id, col = dataset_name)) +
  geom_line(alpha = .3, position = position_jitter(width = .1, height = 0)) + 
  geom_smooth(aes(group = 1)) + 
  scale_y_log10() + 
  geom_hline(lty = 2, yintercept = .5)

```

```{r}
ggplot(filter(d_sub_long, n_trials_rt > 3),
       aes(x = age, y = log_rt, group = subject_id, col = dataset_name)) +
  geom_line(alpha = .3, position = position_jitter(width = .1, height = 0)) + 
  geom_smooth(aes(group = 1)) + 
  scale_y_log10() 
```
# Q1: age-related change in the latent score

Growth in latent language factor. This is the same as the previous factor analysis, but we only look within child and within dataset. 

## 1 factor approach

This has the benefits of simplicity, even though we know that the 1F model is "wrong." 

```{r}
fa_model <-  "F1  =~ rt + rt_var + long_window_accuracy + long_window_acc_var + prod + comp"
fit <- cfa(fa_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

d_sub$f1 <- lavPredict(fit, newdata = d_sub_mat_s)

```

```{r}
ggplot(d_sub |>
         filter(subject_id %in% longitudinal$subject_id), 
       aes(x = age, y = f1, group = subject_id, col = dataset_name)) + 
  geom_point(alpha = .4) + 
  geom_line(alpha = .4) + 
  geom_smooth(aes(group = dataset_name), method = "lm")
```
Note that nesting age slopes doesn't converge. Could go Bayesian if we cared...

```{r}
long_mod <- lmer(f1 ~ age + (1 | subject_id) + (1 | dataset_name), 
                 data = d_sub)

summary(long_mod)
```




## 2 factor approach

Let's try looking at the speed/accuracy model across time.  

```{r}
fa2_model <-  "accuracy =~ long_window_accuracy + long_window_acc_var + comp + prod
               speed =~ rt + rt_var"

fit2 <- cfa(fa2_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

preds <- predict(fit2)

d_sub$accuracy <- preds[,1]
d_sub$speed <- preds[,2]
```

Merge with longitudinal data. 

```{r}
d_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(administration_id) |>
  mutate(admin_num = 1:n(), 
         delta_t = c(0,diff(age)), 
         speed_0 = speed[1]) |>
  group_by(dataset_name) |>
  mutate(speed_0_high = speed_0 > median(speed_0, na.rm=TRUE))
```

Plot. 

```{r}
ggplot(d_long |>
         pivot_longer(speed:accuracy, names_to = "variable", values_to = "value"), 
       aes(x = age, y = value, group = subject_id, col = dataset_name)) + 
  geom_point(alpha = .4) + 
  geom_line(alpha = .4) + 
  geom_smooth(aes(group = dataset_name), method = "lm") + 
  scale_x_log10() +
  facet_wrap(~variable) +
  theme(legend.position = "bottom") + 
  xlab("Age (months)") + 
  ylab("Latent score") 
```
Change in the latents over age. 

```{r}
long_mod_acc <- lmer(accuracy ~ age + (1 | subject_id) + (1 | dataset_name), 
                 data = d_sub)
long_mod_speed <- lmer(speed ~ age + (1 | subject_id) + (1 | dataset_name), 
                 data = d_sub)


summary(long_mod_acc)
summary(long_mod_speed)
```

# Q2: relations between latents and other factors

Longitudinal model. 

```{r}
lmod <- lmer(accuracy ~ age + speed_0 * delta_t + (1 | subject_id) + (1 | dataset_name), 
     data = d_long)

summary(lmod)
```

```{r}
ggplot(d_long, 
       aes(x = age, y = accuracy, group = subject_id, col = dataset_name, lty = speed_0_high)) + 
  geom_line(alpha = .4) + 
  geom_smooth(aes(group = speed_0_high), method = "lm") + 
  facet_wrap(~dataset_name, scales="free_x")
```

